A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2021-06-18 Papers/Paper/manuscript.Rmd
add some citations

keywords          : "TODO"
wordcount         : "X"

@@

```{r setup, include = FALSE}
 library("papaja")
 library(here)
 library(tidyverse)
library(patchwork)
 library(brms)
 library(lme4)
 library(tidybayes)
 library(mgcv)
 library(cowplot)
 library(gridExtra)
library(broom.mixed)
library(tidymv)

@@

Measures of online reading are one way to understand language and how the mind processes language. Many theories of language structure and language processing ground out in predictions about the difficulty of processing words. For instance, the subject v object relative debate includes theories that make fine-grained predictions about which word is how slow -- this needs localized methods to adjudicate it. Other theories such as noisy channel processing also need support from localized word-by-word results. TODO WHY *DO* we care about localization?

Incremental processing methods such as self-paced reading or eye-tracking measure how long someone spends looking at one word before moving on and use that as a proxy for how difficult word was in context. Across multiple methods for measuring per-word processing and reading time, how unexpected a word is correlates with how long it takes to read it and move on (CITE). However, the two major methods of measuring incremental processing both suffer from imprecise localization. In eye-tracking, people read naturally which involves skipping words, jumping ahead and looking back, the dynamics of which make it hard to isolate effects. Even when reading order is controlled in self-paced reading, readers are moving quickly through the text and it may take multiple words for slowdowns to catch up with them.

@@

It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability). Due to spillover effects, on SPR and eye-tracking, there is also a positive linear relationship between the surprisal of a previous word and the RT on the current word -- this is an indication of lack of localization. In addition to suprisal predicting RT, word length and word's overall frequency are also often found to be predictive. TODO many CITES

An alternative method that seems to have superior localization is the Maze task, which adopts an unnatural way of reading to force incremental processing [@forsterMazeTaskMeasuring2009]. In the Maze task, participants see two words at a time, a correct word that continues the sentence, and a distractor which does not. Participants must choose the correct word, and their reaction time (RT) is the dependent measure. If participants make a mistake, the sentence discontinues. Theoretically, participants must fully integrate each word into the sentence in order to confidently select it. This idea is supported by studies finding strongly localized effects [@witzelComparisonsOnlineReading2012a].

@@

### Model specification
To model the relationship between RT and word surprisal, we created a set of predictor variables of frequency, word length, and surprisals from three language models.  For length, we used the length in characters excluding end punctuation. For unigram frequency, we tokenized the training data from @gulordava18 and tallied up instances. We then used the log2 frequency of the expected occurances in 1 billion words as the model predictor, so higher values indicate higher log frequencies. We got per-word surprisals for each of 3 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], and Transformer-XL [@daiTransformerXLAttentiveLanguage2019].  For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.

@@

For linear models, we centered all predictors. We used full mixed effects, including by-subject slopes and a per-word-token random intercept. We used weak priors (normal(1000,1000) for intercept, normal(0,500) for beta and sd, and lkj(1) for correlations). Models were run in BRM.

For model comparison, we fit models with only frequency and length as predictors, as well as models that also had one or more sources of surprisal. We centered all effects. We included a random intercept for word identity, but no further random effects so the models would fit in LME4.

@@

```{r errors, out.height="25%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:error-cap)"}
error_plot <- ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=accurate))+
  geom_point(size=1)+
  labs(x="Fraction words selected correctly",
       y="Mean Reaction Time (ms)")+
  coord_cartesian(xlim=c(0.4,1), ylim=c(0,1600), expand=F)+
    scale_color_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
  guides(color=FALSE)

comp_plot <- ggplot(some, aes(x=num_correct,fill=accurate))+
  geom_bar(position="dodge")+
  facet_grid(.~accurate)+
  labs(x="Comprehension questions correct (out of 6)", y="Participants")+
  scale_fill_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
guides(fill=FALSE)

@@

Participant accuracy reflects both how well participants can navigate the task and what quality the auto-generated distractors are. We calculated the per-word error rate for each participant and graphed it against their average reaction time. (To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded.) As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) make relatively few errors, with some reaching 99% accuracy. This confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story. These careful participants took around 1 second for each word selection, which is much slower than other paradigms CITE. Another cluster of participants (in red) sped through the task, seemingly clicking randomly. This bimodal distribution is likely due to the mix of workers on Mechanical Turk, as we did not use qualification cutoffs.

@@

This extreme incrementality may make the Maze task a good target for comparisons between human RTs and predictions of surprisal from neural language models because spillover effects would not need to be considered as much. 
It shows broadly the same effects as other methods; however the differences in scale of suprisal effects and the lack of frequency effects detected here are a reason to explore more. Comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing.

>>>>

keywords: TODO
wordcount: TODO

@@

```{r setup, include = FALSE}
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))
 library("papaja")
 library(here)
 library(tidyverse)
library(patchwork)
 library(brms)
 library(lme4)
 library(tidybayes)
 library(mgcv)
 library(cowplot)
 library(gridExtra)
library(broom.mixed)
library(tidymv)

@@

Measures of online reading are one way to understand language and how the mind processes language. Many theories of language structure and language processing ground out in predictions about the difficulty of processing words [@bartekSearchOnlineLocality2011]. For instance, the subject v object relative debate includes theories that make fine-grained predictions about which word is how slow -- this needs localized methods to adjudicate it [@grodnerConsequencesSerialNature2005; @staubEyeMovementsProcessing2010; @traxlerProcessingSubjectObject2002]. Other theories such as noisy channel processing also need support from localized word-by-word results [@levyIntegratingSurprisalUncertaininput2011].

Incremental processing methods such as self-paced reading or eye-tracking measure how long someone spends looking at one word before moving on and use that as a proxy for how difficult word was in context. How unexpected a word is correlates with how long it takes to read it and move on [@raynerEffectsFrequencyPredictability2004; @klieglLengthFrequencyPredictability2004]. However, the two major methods of measuring incremental processing both suffer from imprecise localization. In eye-tracking, people read naturally which involves skipping words, jumping ahead and looking back, the dynamics of which make it hard to isolate effects [@raynerEffectsFrequencyPredictability2004; @levyEyeMovementEvidence2009; @frazierMakingCorrectingErrors1982; @raynerEyeMovementsReading1998]. Even when reading order is controlled in self-paced reading, readers may maintain ambiguities about what a word was or means until it is later resolved by context, so it may take multiple words for slowdowns to catch up with them [@macdonaldInteractionLexicalSyntactic1993; @koornneefUseVerbbasedImplicit2006].

@@

It's well established for eye-tracking and SPR that RTs are roughly linear in terms of a word's surprisal (negative log probability) [@smithEffectWordPredictability2013; @wilcoxPredictivePowerNeural]. Due to spillover effects, on SPR and eye-tracking, there is also a positive linear relationship between the surprisal of a previous word and the RT on the current word -- this is an indication of lack of localization. In addition to suprisal predicting RT, word length and word's overall frequency are also often found to be predictive [@klieglLengthFrequencyPredictability2004].

An alternative method that seems to have superior localization is the Maze task, which adopts an unnatural way of reading to force incremental processing [@freedman85; @forsterMazeTaskMeasuring2009]. In the Maze task, participants see two words at a time, a correct word that continues the sentence, and a distractor which does not. Participants must choose the correct word, and their reaction time (RT) is the dependent measure. If participants make a mistake, the sentence discontinues. Theoretically, participants must fully integrate each word into the sentence in order to confidently select it. This idea is supported by studies finding strongly localized effects [@witzelComparisonsOnlineReading2012a].

@@

### Model specification

To model the relationship between RT and word surprisal, we created a set of predictor variables of frequency, word length, and surprisals from three language models.  For length, we used the length in characters excluding end punctuation. For unigram frequency, we tokenized the training data from @gulordava18 and tallied up instances. We then used the log2 frequency of the expected occurances in 1 billion words as the model predictor, so higher values indicate higher log frequencies. We got per-word surprisals for each of 3 different language models: a Kneser-Ney smoothed 5-gram, GRNN [@gulordava18], and Transformer-XL [@daiTransformerXLAttentiveLanguage2019].  For all of these predictors, we consider both the predictor at the current word as well as lagged predictors from the previous word.

@@

For linear models, we centered all predictors. We used full mixed effects, including by-subject slopes and a per-word-token random intercept [@barrRandomEffectsStructure2013]. We used weak priors (normal(1000,1000) for intercept, normal(0,500) for beta and sd, and lkj(1) for correlations). Models were run in brm [@burknerAdvancedBayesianMultilevel2018].

For model comparison, we fit models with only frequency and length as predictors, as well as models that also had one or more sources of surprisal. We centered all effects. We included a random intercept for word identity, but no further random effects so the models would fit in lme4 [@batesFittingLinearMixedeffects2015].

@@

```{r errors, out.height="25%", fig.width=8, fig.height=3, fig.pos='t', fig.cap="(ref:error-cap)"}
error_plot <- ggplot(data_error_summ, aes(x=pct_correct, y=mean_rt, color=accurate))+
  geom_point(size=1)+
  labs(x="Fraction words selected correctly",
       y="Mean Reaction Time (ms)")+
  coord_cartesian(xlim=c(0.4,1), ylim=c(0,1600), expand=F)+
    scale_color_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
  guides(color="none")

comp_plot <- ggplot(some, aes(x=num_correct,fill=accurate))+
  geom_bar(position="dodge")+
  facet_grid(.~accurate)+
  labs(x="Comprehension questions correct (out of 6)", y="Participants")+
  scale_fill_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
guides(fill="none")

@@

Participant accuracy reflects both how well participants can navigate the task and what quality the auto-generated distractors are. We calculated the per-word error rate for each participant and graphed it against their average reaction time. (To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded.) As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) make relatively few errors, with some reaching 99% accuracy. This confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story. These careful participants took around 1 second for each word selection, which is much slower than other paradigms. Another cluster of participants (in red) sped through the task, seemingly clicking randomly. This bimodal distribution is likely due to the mix of workers on Mechanical Turk, as we did not use qualification cutoffs.

@@

This extreme incrementality may make the Maze task a good target for comparisons between human RTs and predictions of surprisal from neural language models because spillover effects would not need to be considered as much [@wilcoxPredictivePowerNeural]. 
It shows broadly the same effects as other methods; however the differences in scale of suprisal effects and the lack of frequency effects detected here are a reason to explore more. Comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing [ex. @bartekSearchOnlineLocality2011].
